### 2019-07-26 paper/paper.tex
Update paper.tex

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% INSERT PSEUDOCODE HERE  FIGURES 2 and 3
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Why GlobalSearchRegression.jl is faster than existing alternatives?}
\label{sec:whygsreg}
\subsection{JULIA platform}
\label{subsub:juliaplat}

Despite some \verb GlobalSearchRegression.jl  specific features to be examined below, our all-subset-regression algorithm benefits from the well-known Julia language efficiency for High Performance Computing tasks.

\vspace{-2mm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% INSERT PSEUDOCODE HERE TABLE 2
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Julia JIT-compilation allows packages to run faster than those executed using interpreted or byte-compiled languages (like R or Stata-Mata). Indeed, basic functions running on Julia can be up to 251 and 368 times faster than those running in R and Stata, respectively -with an average speed-up of 75 and 90, see Table 2-.

\subsection{Parallel strategy and memory setup}
\label{subsub:parallelst}
It is well know that multicore architectures can be used to speed-up execution times through two main paradigms: data and task-parallelism \cite{zaki2001}. The preferred strategy critically depends on:
\begin{enumerate}
    \item Database structure;
    \item Algorithm serial portion; and
    \item Interthread communication costs.
\end{enumerate} 
The first discussion is about tall vs fat data. While tall databases are more suitable for data-parallelism \cite{babu2013}, fat-structures improve relative performance of task-parallelism (because tasks increase exponentially with covariates/columns in feature selection problems, see \cite{foster1994}). \vskip 6pt

Additionally, the choice between alternative paradigms takes into account the Amdahl's Law for the specific algorithm to be used. Task-parallelism is usually better for econometric and machine learning algorithms requiring some specific serial optimization paths (i.e. arima-arfima models), while data-parallelism performs better under linear-algebra solutions (i.e. OLS-family estimators, see \cite{guo2012}). \vskip 6pt

Finally, we have the problem of intercomunication costs. Data-parallelism -generally- involves intense needs of inter-thread-communication. While task-parallelism communication costs depends on Load-Balancing choices (i.e. Dynamic vs Static), they are usually lower than data-parallelism ones \footnote{At least when Static Load Balancing is implemented, because it allows for Coarse-grained granularity \cite{gordon2006}} \cite{MOREANO2017}. \vskip 6pt

For feature selection problems, pros and cons of alternative strategies often determine that available cores should be used for task parallelism. Databases are fat, data-mining algorithms can include large serial portions, and intercomunication costs can be huge for large-multicore architectures. These reasons explain why all available all-subset-regression packages use taks-parallism to speed-up execution times. \vskip 6pt

As for the memory setup, there are also alternative methodologies. First, it's necessary to choose among Static vs. Dynamic memory allocations \citeprajapati2015}. To improve speed-up, Static "once-and-for-all" allocations are often preferred (even at a cost of higher average memory utilization). Second, shared-memory strategies must be determined. Depending on both object-size and CPU architecture -cache size and its distribution among cores-, it could be optimal to use large shared arrays or -alternatively- smaller core-specific objects. Splitting output matrices to work with smaller non-shared arrays could be useful for cache optimization purposes, but it could also entail additional communication costs and higher memory requirements \cite{ahmed2015}. In practice, feature selection algorithms usually prefer shared-arrays for high performance computing. \vskip 6pt

\verb GlobalSearchRegression.jl  (execution-time) advantages have been obtained combining:
\begin{itemize}
    \item[a)]  Task-parallelism
    \item[b)]  Static Load-Balancing
    \item[c)]  Coarse-grained granularity   
    \item[d)]  Static memory allocation
    \item[e)]  Efficient shared-array implementation
\end{itemize}

While some of these characteristics are shared with R and Stata alternatives (\verb MuMin-pdredge  and \verb GSREG,  respectively), our Static Load-Balancing algorithm outperform the round-robin R scheduling (implemented by the clusterapply function included in the \verb parallel  R-package see \cite{deb2013}) and the Shared-Array strategy significantly improves I/O performance against Stata. By construction, pure Static Load-Balancing in GlobalSearchRegression.jl avoids Round-Robin communication costs and execution gaps. This advantage overcompensate minimal\footnote{Minimal because our static scheduler guarantees that average task-complexity will not be too different among workers} load-asymmetries, associated with any static scheduling. In turn, using efficient shared-arrays to store all-subset-regression results allows us to outperform the \verb Stata-GSREG  methodology which heavily relies on slower I/O disk operations (because multiple instances must be launched to enable task-parallelism and, therefore, shared-arrays become unfeasible). \vskip 6pt

\subsection{OLS estimation}
\label{subsub:ols}
Efficient Ordinary Least Squares (OLS) algorithms rely on Linear Algebra operations (matrix decomposition, matrix inversion, etc.).\footnote{Optimization alternatives (i.e. Gradient descent), while quite inefficient for linear models, can also be used for non-linear estimations \cite{abdi2007}} The traditional $(X'X)^{-1}$ operation could be time-expensive with unstable solutions under certain conditions \cite{jameson1981}. A preferred method is the QR-decomposition developed by Francis \cite{francis1961} and Kublanovskaya \cite{kublanovskaya1962s}. The QR factorization allows us to decompose any full rank $N \times p$ matrix $\bar{X}$ as:

\begin{equation}
    X=QR
\end{equation}

where

\begin{itemize}
    \item[•] $Q$ is a $N \times p$ matrix with $Q'Q=I$  ; and
    \item[•] $R$ a $p \times p$ upper triangular matrix.
\end{itemize}

The QR decomposition is fast and provides stable numerical solutions under rank-deficient matrices. Alternative factorizations are either slower (SVD decomposition) or potentially unstable (Cholesky decomposition)\cite{agarwal2014}. \vskip 6pt

\verb GlobalSearchRegression.jl  OLS estimation through QR-decompositions outperform existing Julia alternatives (like \verb GLM.jl)  allowing for large speed-ups compared with \verb R-lm    and \verb Stata-regress  commands. Table 3 shows execution time differences for a 200 covariates - 1000 observations - multivariate linear regression \footnote{Execution times were calculated for alternative OLS algorithms available for all-subset-regression packages in Julia, and Stata.}.
\vskip 6pt

It must be notice that execution times are compilation-free, because all-subset-regression algorithms must perform thousand to millions of regressions where compilation time is absent (it only affects the first regression).
\vskip 6pt
